---
title: "Blog"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
h1.title {font-size: 1px;  color: White;}
body, td {font-size: 18px;}
code.r{font-size: 15px;}
pre {font-size: 12px}
</style>

```{r set up, include=FALSE}
# global options
knitr::opts_chunk$set(fig.width=5, 
                      fig.height=4, 
                      fig.align="center")
```


# Blog

*** 


## 
## Favorite Quotes

> *The best time to plant a tree is twenty years ago. The second-best time is now.* 
>
> <div style="text-align: right"> Dambisa Moyo, Dead Aid </div>

> *而世之奇伟瑰怪非常之观，常在于险远，而人之所罕至焉。故非有志者不能至也。有志矣，不随以止也，然力不足者，亦不能至也。有志与力，而又不随以怠，至于幽暗昏惑而无物以相之，亦不能至也。然力足以至焉，于人为可讥，而在己为有悔；尽吾志也而不能至者，可以无悔矣，其孰能讥之乎？*
>
> <div style="text-align: right"> 王安石（1021-1086）, 《游褒禅山记》</div>
>
> <span style="color:grey">Translation by Xie Baikui</span> 
>
> *So it is regrettable that numerous are excursionists visiting places which are near and easily accessible, but few are travelers who come to places far and dangerous, which, nevertheless, command the most sublime, grotesque and unique views, scarcely enjoyed by ordinary folks. One can never reach such places unless one has a strong will-power. But this alone, though unaffected by easy compliance with others’ inclination to remit one’s efforts, would not ensure the attainment of one’s goal, if not boosted by sufficient strength. A strong will-power and sufficient strength, though unaffected by that sort of easy compliance, would also fail to do so, if unassisted by extrinsic advantages, especially in dark and misleading circumstances. Given sufficient strength, failure to attain one’s objective entails scorn from others and regret from oneself. But, if one does not succeed in doing so after he has given full play to his will-power, he need not feel regret, and who would laugh him to scorn?* 
>
> <div style="text-align: right"> Wang Anshi (1021-1086), A Trip to Baochan Mountain </div>


> *There is only one heroism in the world: to see the world as it is, and to love it.* 
>
> <div style="text-align: right"> Romain Rolland, Life of Michelangelo </div>

> *I liked it, I was good at it, and I was really... I was alive.* 
>
> <div style="text-align: right"> Walter White, Breaking Bad Finale  </div>


***

## Hobbies

- Hiking, squash, table tennis

- Cooking, baking

  Cooking is fun, baking is my therapy

- Chinese calligraphy

  I practiced Chinese calligraph for nine years from 6 to 15. I am proficient in writing regular script (楷书) and clerical script (隶书). During Chinese Spring Festivals, I write Spring couplets for friends and relatives.

- First-/third-person shooter games, e.g. PUBG (吃鸡)

  Strategies matter more than tactics.
  
- Trading index futures
  
  Trading is a dialogue with the world.

- Traveling
  
  I love exploring the world! I have been to 11 countries and regions, and I plan to make it 30 before my 30.


***

## Course Notes 

#### Mathematical Statistics I
![](./pdfs/s301n.pdf){width=100% height=500}


#### Generalized Linear Models
![](./pdfs/s347n.pdf){width=100% height=500}


***

## Stat2Life

I like to applying statistics to life (Stat2Life) so that I can better understand the world and make the right decisions. To share my thinking, I have co-authored a book [The Logic of Doing Things](https://www.amazon.com/Logic-Doing-Things-Always-Chinese/dp/722011172X) with my like-minded university schoolmate Junlin Fu. There are always new ideas coming up, and I briefly summarize some keywords here. If I have time I will add more details.


* Markov Chain, absorbing state, and failure.

* Decision making: maximizing utility v.s. minimizing regret
  c.f. maximizing likelihood v.s. minimizing loss.

* Can one's utilty be a function of other's utility? The derivative can be positive, and can also be negative.


***

## Interesting Problems

During my high school years, one of my brilliant deskmates Yuanke Liu liked to design novel physics problems, and asked me to solve them. Usually these problems involve very basic concepts, but require your rigorous and out-of-the-box thinking. I found this way very helpful to check my understanding, and more importantly, train my ability to integrate basic concepts to solve non-trivial problems. 

When I study statistics and machine learning, I also come up this kind of interesting problems. Below I list a couple of them. Some just check one's understanding of basic concepts, some can be exam questions, and some may look like brain teasers. 

1. In a linear model without intercept, can $R^2$ be larger than 1? Can it be less than 0?

1. Suppose $X_1$ and $X_2$ are two uniform random variables over $[0,1]$. Whether they are dependent or independent is unknown. What is the range of $E\{\max(X_1, X_2)\}$?

1. Write math equations for the forward and backward propagation through a softmax function. This is to check your basic understanding of neural networks.

1. Write Python code for the equations above. Well, this checks your coding skills. And of course, without loops.

1. For a least square problem $\min_{\mathbf{x}\in \mathbb{R}^n} \Vert A \mathbf{x} -\mathbf{b}\Vert_2^2$, find the conditions on $A$ such that the problem has (i) a unique solution; (ii) multiple solutions. Give a sketch of the multiple solutions when $n=2$.

1. Find the derivative of $x^x$. Show as many methods as you can.

1. The basic and classic question: what is $p$-value? Give both mathematical definition and verbal explanation.

1. Explain the logic behind hypothesis testing, verbally and mathematically. In particular, why do we reject the null hypothesis if the $p$-value is small than a threshold?

1. What are type I error and type II error? What is the power of a test?

1. What's the definition of "independent" of two random variables? Give both mathematical definition and verbal explanation.

1. Give an example when $Corr(X_1, X_2)=0$ but $X_1$ and $X_2$ are dependent.

1. What's the relation between maximizing likelihood and minimizing loss? Take a linear model as an example, write the likelihood function and the loss function, and explain their relation.

Solutions to some of the problems:

1. It depends on how you define $R^2$. TBC


1. TBC

1. TBC

1. TBC

1. Suppose there is a solution $\mathbf{x}^*$. If matrix $A$ is not full rank, then there exists $\mathbf{x}_0\in null(A)$ and a scalar $k$ such that $\mathbf{x}=\mathbf{x}^*+k\mathbf{x}_0$ is also a solution, since
$A\mathbf{x}=A\mathbf{x}^*+kA\mathbf{x}_0=A\mathbf{x}^*$. Consider the objective function
$$f(\mathbf{x})=\Vert A \mathbf{x} -\mathbf{b}\Vert_2^2$$
  an example: $$A=\left[\begin{array}{cc}
2 & -2\\
4 & -4
\end{array}\right], \mathbf{b}=\left[\begin{array}{cc}
20 \\
-10 
\end{array}\right].$$
  Clearly, $A$ is singular. The set of solution is $\mathbf{x}=k[1,1]^\top$. This is reflected in the 3D-graph in the sense that the surface of the function $f(\mathbf{x})$ is "folded" along the line $y=x$, as shown below.

```{r message=FALSE}
library(plotly)

f = matrix(0,10,10)
for (x1 in 1:10){
  for (x2 in 1:10){
    f[x1,x2] = (20 - 2*x1 + 2*x2)^2 + (-10 - 4*x1 + 4*x2)^2
  }
}

p <- plot_ly(z = ~f) %>% add_surface(
  contours = list(
    z = list(
      show=TRUE,
      usecolormap=TRUE,
      highlightcolor="#ff0000",
      project=list(z=TRUE)
      )
    )
  ) %>%
  layout(
    scene = list(
      camera=list(
        eye = list(x=2, y=2, z=2)
        )
      )
  )

p

```




